---
header-includes:
   - \usepackage{bbm}
always_allow_html: yes
title: "Results"

bookdown::html_document2: default
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: true
      smooth_scroll: true
    theme: yeti
    highlight: default
---

&nbsp;
&nbsp;
&nbsp;
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SVM

We used Support Vector Classifiers with Radial Basis function as kernel. With the hyper-parameter tuning, we found value of cost = 1 and gamma = 0.5 to be optimal w.r.t our ground truth.


### Library Imports

The CRAN packages `e1071` for classification using Support Vector Classifiers with Radial Basis Function kernel and hyper-parameter tuning and `ROSE` for over-sampling of data are being used.
```{r library, echo=TRUE}
library(e1071)
library(ROSE)
```


### Pre-Processing
```{r preprocessing, echo=FALSE}

load('covid_data_new_masked')
dataset <- covid_data
summary(dataset)


#Factoring the Target variable of (0,1)
dataset$V59 = factor(dataset$V59, levels = c(0, 1))
print(is.factor(dataset$V59))

#Dataset Split
index <- 1:nrow(dataset)
testindex <- sample(index, trunc(length(index)/3))
testset <- dataset[testindex,]
trainset <- dataset[-testindex,]


#Over-Sampling
train_ov_sampled <- ovun.sample(V59 ~ ., trainset, method = "over")$data
table(train_ov_sampled$V59)

test_ov_sampled <- ovun.sample(V59 ~ ., testset, method = "over")$data
table(test_ov_sampled$V59)

```

### Classification
``` {r classification, echo=TRUE}
set.seed(825)

#fitting the Support Vector Machine to the Training set
svm_fit_ovs <- svm(V59~., kernel = 'radial',
               data = train_ov_sampled, scale=TRUE, cachesize = 200,
               probability = TRUE, gamma = 0.5, cost = 1)

summary(svm_fit_ovs)


#Hyper-Parameter Tuning
# obj <- tune.svm(V59~., data = train_ov_sampled, sampling = "cross",
#           gamma = 2^c(-1:4), cost = c(1:10))
# summary(obj)
```

### Prediction
```{r prediction, echo=FALSE}
#Predicting the test set results
svm.pred.ovs <- predict(svm_fit_ovs, test_ov_sampled[,-59], probability = TRUE)


plot(svm.pred.ovs)

summary(svm.pred.ovs)
```


### Evaluation
```{r evaluation, echo=FALSE}
#Evaluation on Test_Set
accuracy.meas(response = test_ov_sampled$V59, predicted = svm.pred.ovs)

roc.curve(test_ov_sampled$V59, svm.pred.ovs, plotit = T)

caret::confusionMatrix(svm.pred.ovs, test_ov_sampled$V59, positive = '1')
```

