---
header-includes:
   - \usepackage{bbm}
always_allow_html: yes
title: "Results"

bookdown::html_document2: default
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: true
      smooth_scroll: true
    theme: yeti
    highlight: default

link-citations: yes
bibliography: references.bib
csl: data-and-knowledge-engineering.csl
---

&nbsp;
&nbsp;
&nbsp;



## Benchmark Evaluation {.tabset}

```{r accuracyplotmodel, echo=FALSE}


x<- c("KNN", "Logistic Regresssion", "Random Forests", "Decision Tree", "Naive Bayes", "SVM")
y<- c(85.27, 90.34, 87.84, 87.23, 83.84, 54.23)

x_name <- "Models"
y_name <- "Accuracy"

df <- data.frame(x,y)
names(df) <- c(x_name,y_name)

library(ggplot2)
ggplot(data=df, aes(x=Models, y=Accuracy)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Accuracy), vjust=-0.3, size=3.5)+
  ggtitle("Accuracy of models on raw dataset")+
  theme_light()


```
 
In accordance to [@ozturk2020automated], where in the author presented a network for detection of COVID-19  cases using Deep Neural Networks with X-Ray images.The paper presents a detailed comparison study of their network with respect to other models for COVID-19 detection, in terms of accuracy. The details of the data-set used is explained in the following paragraph.
 
 
The database was developed by [@cohen2020covid] using images from various sources. The database is constantly upgraded, as of now, the database content comprise of around "201" COVID-19 positive X-Ray images.There were no images, which represent  X-Ray of normal lungs, hence we have used thousand healthy X-Ray images from our original database as negative class, which we have used to feed our original models.
 
In the following sections, we present the figures that are obtained by training the models with the same data-set that [@ozturk2020automated] used in his experiment followed by the comparison table :
 
#### <span class="sub-header">Comparison of different models</span>


| Study           | Type Of Image | No. Of Cases                     | Method Used and Settin              | Accuracy |
|-----------------|---------------|----------------------------------|-------------------------------------|----------|
| Sethy and Behra | Chest X-ray   | 25 COVID-19(+) 25 COVID-19 (−)   | ResNet50+ SVM                       | 95.38    |
| Hemdan et al.   | Chest X-ray   | 25 COVID-19(+) 25 Normal         | COVIDX-Net                          | 90.0     |
| Narin et al.    | Chest X-ray   | 50 COVID-19(+) 50 COVID-19 (−)   | Deep CNN ResNet-50                  | 98.0     |
| Ying et al.     | Chest CT      | 777 COVID-19(+) 708 Healthy      | DRE-Net                             | 86.0     |
| Wang et al.     | Chest CT      | 195 COVID-19(+) 258 COVID-19(−)  | M-Inception                         | 82.9     |
| Zheng et al.    | Chest CT      | 313 COVID-19(+) 229 COVID-19(−)  | UNet+3D Deep Network                | 90.8     |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | Decision Tree with RAW Data         | 90.03    |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | Decision Tree with Oversampled Data | 69.05    |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | Random Forest with RAW Data         | 91.69    |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | Random Forest with Oversampled Data | 79.17    |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | kNN with RAW Data                   | 90.62    |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | Logistic Regression with RAW Data   | 90.34    |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | Naive Bayes with RAW Data           | 87.0     |
| Our Study       | Chest X-ray   | 1000 COVID-19(-) 201 COVID-19(+) | SVM with Oversampled Data           | 54.73    |

***

### Random Forest 

```{r random forest , echo =TRUE }

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction   0   1
#          0 245  18
#          1   7  31
#                                           
#               Accuracy : 0.9169          
#                  95% CI : (0.8798, 0.9455)
#     No Information Rate : 0.8372          
#     P-Value [Acc > NIR] : 3.802e-05       
#                                           
#                   Kappa : 0.665           
#                                           
#  Mcnemar's Test P-Value : 0.0455          
#                                           
#             Sensitivity : 0.6327          
#             Specificity : 0.9722          
#          Pos Pred Value : 0.8158          
#          Neg Pred Value : 0.9316          
#               Precision : 0.8158          
#                  Recall : 0.6327          
#                      F1 : 0.7126          
#              Prevalence : 0.1628          
#          Detection Rate : 0.1030          
#    Detection Prevalence : 0.1262          
#       Balanced Accuracy : 0.8024          
#                                           
#        'Positive' Class : 1



```


### kNN

```{r knn CM, echo =TRUE}


# Confusion Matrix and Statistics
# 
#           Reference
# Prediction   0   1
#          0 283  21
#          1  12  36
#                                           
#                Accuracy : 0.9062          
#                  95% CI : (0.8709, 0.9346)
#     No Information Rate : 0.8381          
#     P-Value [Acc > NIR] : 0.0001493       
#                                           
#                   Kappa : 0.6311          
#                                           
#  Mcnemar's Test P-Value : 0.1637344       
#                                           
#             Sensitivity : 0.9593          
#             Specificity : 0.6316          
#          Pos Pred Value : 0.9309          
#          Neg Pred Value : 0.7500          
#              Prevalence : 0.8381          
#          Detection Rate : 0.8040          
#    Detection Prevalence : 0.8636          
#       Balanced Accuracy : 0.7955          
#                                           
#        'Positive' Class : 0


```



### Logistic Regression
```{r Logistic Regression CM, echo =TRUE}

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction   0   1
#          0 282  21
#          1  13  36
#                                           
#                Accuracy : 0.9034          
#                  95% CI : (0.8676, 0.9322)
#     No Information Rate : 0.8381          
#     P-Value [Acc > NIR] : 0.0002803       
#                                           
#                   Kappa : 0.6228          
#                                           
#  Mcnemar's Test P-Value : 0.2299491       
#                                           
#             Sensitivity : 0.9559          
#             Specificity : 0.6316          
#          Pos Pred Value : 0.9307          
#          Neg Pred Value : 0.7347          
#              Prevalence : 0.8381          
#          Detection Rate : 0.8011          
#    Detection Prevalence : 0.8608          
#       Balanced Accuracy : 0.7938          
#                                           
#        'Positive' Class : 0    
```



### Decision Tree

```{r decision tree, echo =TRUE }
# Confusion Matrix and Statistics
# 
#           Reference
# Prediction   0   1
#          0 243  21
#          1   9  28
#                                           
#                Accuracy : 0.9003          
#                  95% CI : (0.8608, 0.9317)
#     No Information Rate : 0.8372          
#     P-Value [Acc > NIR] : 0.001154        
#                                           
#                   Kappa : 0.5943          
#                                           
#  Mcnemar's Test P-Value : 0.044610        
#                                           
#             Sensitivity : 0.57143         
#             Specificity : 0.96429         
#          Pos Pred Value : 0.75676         
#          Neg Pred Value : 0.92045         
#               Precision : 0.75676         
#                  Recall : 0.57143         
#                      F1 : 0.65116         
#              Prevalence : 0.16279         
#          Detection Rate : 0.09302         
#    Detection Prevalence : 0.12292         
#       Balanced Accuracy : 0.76786         
#                                           
#        'Positive' Class : 1


```




### Naive Bayes

```{r navie bayes , echo =TRUE }
## Confusion Matrix and Statistics
## 
##           Reference
## Prediction Normal COVID19
##    Normal     235      24
##    COVID19     15      26
##                                           
##                Accuracy : 0.87            
##                  95% CI : (0.8266, 0.9059)
##     No Information Rate : 0.8333          
##     P-Value [Acc > NIR] : 0.04857         
##                                           
##                   Kappa : 0.4957          
##                                           
##  Mcnemar's Test P-Value : 0.20018         
##                                           
##             Sensitivity : 0.52000         
##             Specificity : 0.94000         
##          Pos Pred Value : 0.63415         
##          Neg Pred Value : 0.90734         
##              Prevalence : 0.16667         
##          Detection Rate : 0.08667         
##    Detection Prevalence : 0.13667         
##       Balanced Accuracy : 0.73000         
##                                           
##        'Positive' Class : COVID19         
## 


```



### Support Vector Machine

```{r svm, echo =TRUE }

# 
# Confusion Matrix and Statistics
# 
#           Reference
# Prediction   0   1
#          0  26   0
#          1 307 346
#                                           
#                Accuracy : 0.5479          
#                  95% CI : (0.5096, 0.5858)
#     No Information Rate : 0.5096          
#     P-Value [Acc > NIR] : 0.02506         
#                                           
#                   Kappa : 0.0795          
#                                           
#  Mcnemar's Test P-Value : < 2e-16         
#                                           
#             Sensitivity : 1.00000         
#             Specificity : 0.07808         
#          Pos Pred Value : 0.52986         
#          Neg Pred Value : 1.00000         
#              Prevalence : 0.50957         
#          Detection Rate : 0.50957         
#    Detection Prevalence : 0.96171         
#       Balanced Accuracy : 0.53904         
#                                           
#        'Positive' Class : 1

```

***

## Final Analysis

Based on our approach, we initially tried to extract the features out of the CXR images without any segmentation. However, we found that the area of interest i.e., the lung area is a very small segment of the whole image. Hence, the noise around the lung area dominates the outcome of the classification. Moreover, we found that quality of CXR images highly affects the feature vectors. For this reason, we segmented the images to find the area of interest and extract our features on corresponding lung-segment. The results were better in that case. However, we could not deal with quality of images in terms of light exposure. 

We tried explaining our prediction using `LIME` image module. Intially, we tried applying LIME on the whole image. However, the results were not good. All the areas highlighted by LIME were outside of lung-portion. So, we tried applying LIME on segmented image by applying black background on the non-lung area hoping that the LIME super-pixels will ignore the black backgrounded regions. We manually confirmed that any perturbed image outside the lung-segment area doesn't get a different prediction. However, we partially succeded on this approach where we noticed that only the negative COVID images (False Negatives too) were correctly marked by LIME. However, positive images could not be properly annotated by LIME within the area of interest.

We also applied classification algorithms, on benchmark dataset and compared the performaces of simple intrinsically explainable models we used and found that some of the very simple models such as kNN, Naive Bayes and Random Forest are giving high accuracy. To our astonishment, SVM couldn't perform well in comparsion to these simpler models. 

We proposed to work on an ensemble architecture where we use our simpler models to boost our confidence in prediction. However, such an approach couldn't give us the kind of robustness we expected. Hence, after seeing results with the ensemble approach, we dropped the idea of ensemble classification.

We found Local Binary Pattern to be a powerful texture based descriptor, where the only justification of its usefulness can be seen when we compare our results with Neural Network based approaches in the benchmark section.

***

## Conclusion
We applied intrinsically explainable models and LIME explainer on CXR images, for COVID detection and explaination. The results obtained by the classifiers were of higher accuracy whereas LIME failed to perform the similarly. We observed that LBP are good dicriminatory feature vectors for CXR images, however we cannot mitigate the effect of luminosity, brightness and quality to the images. These factors add bias to our classifiers.

***

## References